Query Optimization for Dynamic Imputation

### 2017-05-11 paper/algorithm.tex
Add mean and hot-deck descriptions

\subsubsection{Query runtime}
To compute $\textsc{Time}_{T}(q)$, an additive measure of runtime cost, we retain a simple set of heuristics used by a standard database.

@@

\ProjectName{} is designed so that any imputation strategy
can be plugged in with minimal effort and without changes to the optimizer, simply by supplying
penalty and time functions (\Cref{sec:cost-model}) pertinent to the algorithm.
These can then be queried by the optimizer during planning.  The flexibility of
this approach aligns with the spirit of query optimization, which aims to
reconcile performance with a declarative interface to data manipulation.

@@

\[
    P(q) = \frac{1}{\sqrt{ \textsc{Attr}(q) * \textsc{Card}(q)}}
\]

@@

\todo{Fill this in.}
\[P(q) = \frac{1}{n} (1 + \overline{Variance(q)}) \]

@@

\todo{Fill this in.}

>>>>

\subsubsection{Query runtime}
\label{sec:query-runtime}

To compute $\textsc{Time}_{T}(q)$, an additive measure of runtime cost, we retain a simple set of heuristics used by a standard database.

@@

\ProjectName{} is designed so that any imputation strategy
can be plugged in with minimal effort and without changes to the optimizer, simply by supplying
penalty and time functions (\Cref{sec:cost-model}) pertinent to the algorithm.
These can then be queried by the optimizer during planning.  The flexibility of
this approach aligns with the spirit of query optimization, which aims to
reconcile performance with a declarative interface to data manipulation. To
motivate this API, we describe how an analyst would specify the form of these
functions for a general-purpose imputation algorithm, as well as two simpler
methods.

@@

\[
    P(q) = \frac{|\textsc{Dirty}(q)|}{\sqrt{ \textsc{Attr}(q) * \textsc{Card}(q)}}
\]

@@

At the other end of the spectrum of complexity, an analyst could use a simpler
method that replaces a missing value with the mean of the complete values of
that attribute. Mean value imputation is a common choice for use in exploratory
data analysis. One benefit of mean imputation in the context of \ProjectName{}
is that taking advantage of the histograms associated with a plan $q$, we can
estimate the mean and impute missing values in a non-blocking fashion. Thus,
applying the heuristics from \ref{sec:quality} and \ref{sec:query-runtime}, we
have

\begin{equation*}
\begin{aligned}
    T(q) &\approx 0 \\
    P(q) &= \frac{|\textsc{Dirty}(q)|}{\sqrt{\textsc{Card}(q)}} \\
    P(q) &= \frac{1}{n} (1 + \overline{Variance(q)})
\end{aligned}
\end{equation*}
>>>>>>> Add mean and hot-deck descriptions

@@

Hot-deck imputation is a nonparametric method in which missing values are
replaced with randomly selected complete values. This method preserves the
distribution of the complete values. The heuristic $T$ and $P$ functions are
similar except for the fact that hot-deck is blocking in our implementation.

\begin{equation*}
\begin{aligned}
    T(q) &= \textsc{Card}(q) \\
    P(q) &= \frac{|\textsc{Dirty}(q)|}{\sqrt{\textsc{Card}(q)}}
\end{aligned}
\end{equation*}
